Characterizing Stellar Streams by ML

### 2023-08-28 src/tex/ms.tex
[showyourwork] overleaf sync

    In this section we apply our model to a synethic stream in order to
    demonstrate an accurate recovery of stream members and density variations.

    We generate synthetic observations by sampling points from a third degree
    polynomial \JN{check this} with a constant scatter. Density variations are
    then simulated by evaluating the stream density with two Gaussian
    distributions centered at different values of $\phi_1$ with random
    amplitudes (which sum to $1$) and widths. Stream stars are then removed
    using random sampling, weighted by their likelihood of belonging to the
    density. The background astrometric  points in are generated from
    exponential distributions at each $\phi_1$. The number of stream stars is
    $\variable{output/mock/nstream_variable.txt}$, while the number of
    background stars is $\variable{output/mock/nbackground_variable.txt}$.

@@

    The synthetic stream is also simulated in magnitudes using a 12~Gyr
    isochrone with [Fe/H] = -1.35~dex. The isochrone truncates just short of the
    giant branch, and sampled from uniformly (i.e., the initial mass function,
    $P(\gamma)$, is uniform). The background photometric coordinates are
    generated as a 2-dimensional Gaussian with non-zero covariance. The distance
    modulus is derived from a polynomial parallax track with uniform thickness.
    Thus, the synthetic data scenario described here generates a mock stream in
    astrometric coordinates and photometric coordinates, with density
    variations, a distance gradient which is consistent in parallax and distance
    modulus (i.e., magnitudes), all superimposed over a noisy field of
    background points in each astrometric and photometric dimension.

>>>>

    We generate synthetic stream observations by sampling uniformly in $\phi_1$
    In $\phi_2$ we sample from a quadratic in $\phi_1$ with normally-distributed
    $0.15\degree$ scatter.  The distances are sampled from a linear funciton in
    $\phi_1$ with normally-distributed $0.25 \unit{kpc}$ scatter, then
    transformed to the parallax $\parallax$.  Non-poissonian density variations
    are then simulated by evaluating the stream density with two Gaussian
    distributions centered at different values of $\phi_1$ with random
    amplitudes (which sum to $1$) and widths.  Stream stars are then removed
    using random sampling, weighted by their likelihood of belonging to the
    density. The background astrometric points in are generated from exponential
    distributions at each $\phi_1$.  The number of stream stars is
    $\variable{output/mock/nstream_variable.txt}$, while the number of
    background stars is $\variable{output/mock/nbackground_variable.txt}$.

    The synthetic stream's photometry is also simulated using a 12~Gyr, [Fe/H] =
    -1.35~dex NMIST isochrone \TODO{[cite brutus]}. The isochrone truncates just
    short of the giant branch, and is sampled from uniformly along the isochrone
    (i.e., the initial mass function, $P(\gamma)$, is uniform) as well as with
    an intrinsic uniform 0.2~dex scatter orthogonal to the isochrone track. The
    isochrone is shifted by the stream's parallax modulus, which is derived from
    the astrometric distance track.  The background photometric coordinates are
    generated as a 2-dimensional Gaussian with non-zero covariance.  Thus, the
    synthetic data scenario described here generates a mock stream in
    astrometric coordinates and photometric coordinates, with density
    variations, a distance gradient which is consistent in parallax and distance
    modulus (i.e., magnitudes), all superimposed over a noisy field of
    background points in each astrometric and photometric dimension.

@@

    We pre-train a normalizing flow (see
    \autoref{sub:method:photometric_model:off_stream}) on an off-stream
    selection.  We train on the photometry.  You can see the selection in
    \auroref{fig:mock_data_photometric_background_selection}.  You can see both
    the on-stream and off-stream labels, blue and black respectively.  We select
    using an astrometric selection. In the photometry you can see that the
    on-stream selection has two populations, drawn from the isochrone and drawn
    from the background The off-stream selection, meanwhile, has only the
    background population. Next we train the full model.
